\Problems{exerCatMap}{7mar2018}
% siminos/spatiotemp/Problems/exerCatMap.tex called by catMap.tex
% $Author: predrag $ $Date: 2021-06-12 01:01:43 -0400 (Sat, 12 Jun 2021) $

% Predrag                                               23jan2018

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\exercise{Cat map Green's function, infinite lattice.}{\label{exer:catMapGreenInf}

(a) Show that the eigenvalues of the cat map $M$ are
given by
\beq
\ExpaEig^{\pm}=\frac{1}{2}(s\pm \surd{D})
\,,\qquad
\ExpaEig=e^{\Lyap}
\,,
\ee{catEigsEx}
where $\ExpaEig\equiv\ExpaEig^+$, $s=\ExpaEig+\ExpaEig^{-1}$,
$\surd{D}=\ExpaEig-\ExpaEig^{-1}$, and the discriminant is $D={s^{2}-4}$.

(b) Verify by substitution that the Green's function is given by
\beq
\gd_{nn'}= \frac{1}{{\surd{D}}}\,\frac{1}{\ExpaEig^{|n'-n|}}
\,.
\ee{GreenFun00b}

(c) Show that the orbit is then recovered by
\beq
\ssp_{n} =  \frac{1}{{\surd{D}}}
\sum_{n' \in \integers} \ExpaEig^{- |n-n'|} \Ssym{n'}
\,.
\label{BirVivx=s}
\eeq
    } % end \exercise{exer:catMapGreenInf}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\exercise{Cat map Green's function for a \po.}{\label{exer:catMapGreenPBC}
Show that the Green's function for a \po\ of period $\cl{p}$ is
obtained by summing \refeq{GreenFun00b} over period \cl{p}:
\beq
g^p_{nn'}=\sum_{j=-\infty}^{\infty}{g}_{n-n',j\cl{p}}
  =\frac{1}{{\surd{D}}}\,\frac{\ExpaEig^{-|n-n'|}+\ExpaEig^{-\cl{p}+|n-n'|}}
        {1-\ExpaEig^{-\cl{p}}}
\,.
\ee{GreenFun2}
Verify this formula by explicit matrix inversion for a few periodic points
of cycles $p$ of
periods $\cl{p}=1,2,3,4,\cdots$.
    } % end \exercise{exer:catMapGreenPBC}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\exercise{$d=2$ cat map guess Green's function, infinite lattice.}{\label{exer:catMapGreend=2wrong}
% Predrag 2018-03-11 wrong guess, debunked by Han
Show by substitution that a $d=2$ ``Green's function'' guess given by
\beq
\gd_{z z'}= \PCedit{\frac{1}{2}}\frac{1}
            {{\surd{D}}}\,\frac{1}{\ExpaEig^{|\ell'-\ell|+|t'-t|}}
\,,
\ee{GreenFund=2wrong}
(and similarly, in arbitrary dimension $d>1$) \emph{does not} satisfy the
Green's function conditions
\beq
(\D \gd)_{zz'}=\delta_{zz'}=\delta_{ll'}\delta_{tt'}
\,,
\ee{HLGreenFund=2C}
Here the eigenvalues of the cat map $M$ are
\beq
\ExpaEig^{\pm}=\frac{1}{2}(s/2 \pm \surd{D})
\,,\qquad
\ExpaEig=e^{\Lyap}
\,,
\ee{catEigsEx2D}
where $\ExpaEig\equiv\ExpaEig^+$, $s/2=\ExpaEig+\ExpaEig^{-1}$,
$\surd{D}=\ExpaEig-\ExpaEig^{-1}$, and the discriminant is $D={(s/2)^{2}-4}$.

Hint: the check works just like for \refexer{exer:catMapGreenInf}.
    } % end \exercise{exer:catMapGreend=2wrong}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\exercise{\Po s of Arnol'd cat map.}{\label{exer:AKScatMapPOs}
% moved to here from {blogAKS}{2016-06-09}
\begin{enumerate}
  \item %[[X]]
Describe precisely how you actually pick ``random $q_1$ and $q_2$''
  \item %[[X]]
Explain what happens if $q_1$ and $q_2$ are rational
  \item %[[X]]
Can you get a \po\ if $q_1$ and $q_2$ are irrational?
  \item %[[X]]
What do you mean by period $0$?
  \item %[[X]]
Does the Arnol'd cat map have \po s of any period?
  \item %[[X]]
Derive analytically that $m_j \in \{-1, 0,  1,2\}$ (you can continue the
exposition that I started in \refsect{sect:PerViv}, if that helps). Does you
result agree with Percival and Vivaldi\rf{PerViv}?
\end{enumerate}
%                                       \PCpost{2016-06-02}{:
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\exercise{The second iterate generating partition.}{\label{exer:n=2GenPart}
%    \PCpost{2019-08-10}{
\refFig{fig:PVAdlerWeissS} is very helpful in giving us a visual
understanding of what a Hamiltonian cat map does, and how the generating
partition comes about. Draw the corresponding \AW\ generating partition
for $s=3$, the second, $n=2$ iterate, to verify that the $n=1$ determines
a generating partition for all subsequent times.
    } % end \exercise{exer:n=2GenPart}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\exercise{XXX.}{\label{exer:XXX}
%XXX
%    } % end \exercise{exer:XXX}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\exercise{XXX.}{\label{exer:XXX}
%XXX
%    } % end \exercise{exer:XXX}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \ProblemsEnd
