% siminos/spatiotemp/Examples/examHill2ndOrd.tex
% $Author: predrag $ $Date: 2021-12-24 01:25:20 -0500 (Fri, 24 Dec 2021) $
% Predrag 2020-12-14 called by siminos/spatiotemp/stability.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \ifblog
\subsection{Hill's formula for a second-order system}
    \else % called by kittens/CL18.tex
\subsubsection{Hill's formula for a second-order system.}
    \fi
\label{exam:Hill2ndOrd}
% based on siminos/spatiotemp/Examples/examStab3cyc.tex 2020-12-14

\PC{2020-12-15}{
Merge \refsect{sect:catlattHill}~{\em \catLatt\ Hill's formula}
into his, to transfer to CL18.tex. Once incorporated, remove from here
(too much duplication, as is)
    }

Consider an period-$\cl{}$ {\lattstate} $\Xx_p$, with $d$ fields
$\{\ssp_{\zeit,1},\ssp_{\zeit,2},\dots,\ssp_{\zeit,d}\}$ on each temporal
lattice site $\zeit$ satisfying the condition
\beq
\ssp_{\zeit+1} - \flow{}{\ssp_{\zeit-1}} + \ssp_{\zeit-1} = 0
    \,,\quad
\zeit=1,2,\cdots,\cl{}
\,,
\ee{2stepNonlimTemp}
where $\flow{}{\ssp}$ is a $d$\dmn\ function.

Consider one-dimensional Schrödinger
equation\rf{Peierls33,Harper55,Hofstadter76,Simon82},
$H = -d^2/dx^2 + V(x)$ with $V(x)$
almost periodic and the discrete (= tight binding) analog, i.e., the
doubly infinite Jacobi matrix
\beq
h_{ij} = \delta_{i,j+1}  + V_i\delta_{i,j} + \delta_{i,j-1}
\ee{Simon82a}
with $V_n$ almost periodic on the integers.

A deviation $\Delta\Xx$ from $\Xx_p$ then satisfies the linearized condition
\beq
\Delta\ssp_{\zeit} - \jMat_{\zeit-1}\,\Delta\ssp_{\zeit-1} = 0
\,,\qquad
(\jMat_{\zeit})_{ij}
=
\left.\frac{\partial \flow{}{\ssp}_i}
           {\partial \ssp_{j}}\right|_{\ssp_{j}=\ssp_{\zeit,j}}
\,,
\ee{d-1stepJac3}
where $\jMat_{\zeit}$ is the two-configuration $[d\!\times\!{d}]$
\jacobianM.

in our notation
\refeq{JiKoKr20(2)} is written as
\refeq{eq:CatMapNewton5}
\beq
\ssp_{\ell+1}  -  {s}_\ell\, \ssp_{\ell} + \ssp_{\ell-1}
    =
-\Ssym{\ell}
\,,
\ee{Harper1}

He, as everybody else, sooner or later, rewrites \refeq{Harper} in
the ``\PV\'' `two-configuration representation'\rf{Hofstadter76,PerViv}
matrix $\jMps$
%=\MatrixII{0}{1}
%          {-1}{s}
\refeq{PerViv:2confRepMat},
\beq
 \left[\begin{array}{c}
 \Delta\field_{\zeit}  \\
 \Delta\field_{\zeit+1}
 \end{array} \right]=
 \left[\begin{array}{cc}
  0     & \id_d       \\
 -\id_d & \jMat_{\zeit}
 \end{array} \right]
 \left[\begin{array}{c}
 \Delta\field_{\zeit-1}  \\
 \Delta\field_{\zeit}
 \end{array} \right] %\mbox{ mod } 1
\,.
\ee{Hill:StateSpCatMap}
(note `upside-down' 2D vector), and concerns himself with
the rational,
energy $\tr \jMps^m<2$, oscillatory case.

Note, it's only about the order of recurrence, nobody says the systems
should be Hamiltonian / Lagrangian, works for dissipative systems as
well.





A deviation $\Delta\Xx$ from $\Xx_p$ then satisfies the linearized condition
\beq
\Delta\ssp_{\zeit} - \jMat_{\zeit-1}\,\Delta\ssp_{\zeit-1} = 0
\,,\qquad
(\jMat_{\zeit})_{ij}
=
\left.\frac{\partial \flow{}{\ssp}_i}
           {\partial \ssp_{j}}\right|_{\ssp_{j}=\ssp_{\zeit,j}}
\,,
\ee{d-1stepJac4}
where $\jMat_{\zeit}$ is the two-configuration $[d\!\times\!{d}]$
\jacobianM.

It suffices to work out a temporal period $\cl{}=3$ example to
understand the calculation for any period.
In terms of the $[3d\!\times\!3d]$ matrix shift matrix $\shift$,
the \jacobianOrb\ % \refeq{fcnewt}
can be written as
\beq
\jMorb_p \,=\,
\id - \shift^{-1} \jMat
\,,\quad
\shift^{-1} =
\left[
\begin{array}{ccc}
0     & 0     & \id_d  \\
\id_d & 0     & 0  \\
0     & \id_d & 0
\end{array}
\right]
\,,\quad
\jMat =
\left[
\begin{array}{ccc}
\jMat_1 & 0 & 0 \\
0 & \jMat_2 & 0  \\
0 & 0 & \jMat_3
\end{array}
\right]
\,,
\ee{3shift2}
where $\id_d$ is the $d$\dmn\ identity matrix.
Next, note that
\bea
(\shift^{-1}\jMat)^2 &=&
\shift^{-2}
\left[
\begin{array}{ccc}
\jMat_2\jMat_1 & 0 & 0 \\
0 & \jMat_3 \jMat_2 & 0  \\
0 & 0 & \jMat_1\jMat_3
\end{array}
\right]
,\;\;
(\shift^{-1}\jMat)^3  =
\left[
\begin{array}{ccc}
\jMat_2\jMat_1\jMat_3 & 0 & 0 \\
0 & \jMat_3\jMat_2\jMat_1 & 0  \\
0 & 0 & \jMat_1\jMat_3\jMat_2
\end{array}
\right],
\nnu %\label{stabSquare}
\eea
as $\shift^{-3}=\id$.
Likewise, as $\shift^\cl{}=\id$ for any period $\cl{}$, the trace of
$[\cl{}d\!\times\!\cl{}d]$ matrix
\[
\Tr(\shift^{-1}\jMat)^k=\delta_{k,r\cl{}}\,\cl{}\,\tr\jMat_p^r
\,,\quad
\jMat_p = \jMat_\cl{}\jMat_{\cl{}-1}\cdots\jMat_2\jMat_1
\]
is non-vanishing only if $k$ is a multiple of $\cl{}$, with $\jMat_p$ the
forward-in-time $[d\!\times\!{d}]$ \jacobianM\ of the \po\ $p$.

Now we can evaluate the {\HillDet} \refeq{eq:fundFact}
by expanding
\bea
\ln\Det(\jMorb_p) &=&
\Tr\ln(\id-{\shift}^{-1}\jMat)
                \,=\,
-\sum_{k=1}^\infty\frac{1}{k}\,\Tr({\shift}^{-1}\jMat)^k
    \continue
                 &=&
-\tr\sum_{r=1}^\infty\frac{1}{r} \jMat_p^{r}
  =
\ln\det(\id_d-\jMat_p)
\,.
\label{LnDet=TrLn3}
\eea
The \HillDet\ for a 2-term difference equation \refeq{1stepNonlimTemp} on a
temporal lattice
\[
\Det(\jMorb_p) = \det(\id_d-\jMat_p)
\] %merge with CL18 equation
thus relates the global orbit stability to the Floquet, temporal
evolution stability.
In the {temporal Bernoulli} case, the field $\ssp_\zeit$ is a scalar, and
the 1-time step $[1\!\times\!1]$ time-evolution \jacobianM\
\refeq{d-1stepJac2} at every lattice point $\zeit$ is simply
$\jMat_{\zeit}={s}$, so
\beq
N_\cl{} = |\Det\jMorb| = {s}^{\cl{}} - 1
\,,
\ee{detBern2n3}
in agreement with the time-evolution count. % \refeq{noPerPtsBm}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
