% siminos/spatiotemp/chapter/blogXW.tex
% $Author: predrag $ $Date: 2021-12-22 18:29:33 -0500 (Wed, 22 Dec 2021) $

% called by siminos/spatiotemp/blogCats.tex


\chapter{Xuanqi's blog}
\label{chap:blogXW}
% Predrag                                       2021-12-07
\renewcommand{\Ssym}[1]{{\ensuremath{m_{#1}}}}    % Boris
\renewcommand{\Refl}{\ensuremath{{\sigma}}} % Dihedral wiki convention
\renewcommand{\shift}{\ensuremath{r}}
\renewcommand{\ssp}{\ensuremath{\phi}}             % lattice site field

	
{\large Xuanqi Wang}\\
xwang3021@gatech.edu\\
subversion siminos : xwang3021 \\ % spatioTemp \\
cell +86 177-2115-8435 \\
\hfill   WeChat  y2528742620

   % *********************************************************************
\hfill   {\color{red} The latest entry at the bottom for this blog}
\bigskip

\bigskip

\begin{description}

\item[2021-12-07 Predrag] to Xuanqi:

As you go along, write up your narrative in this file, ask questions -
this is your personal blog, like an experimentalist's log - everything
that you learn and want to share goes in here.
Clip \& paste anything from other sections you
want to discuss, that saves you LaTeXing time.

\item[2021-09-09 Predrag]
The 3rd line of  \emph{siminos/spatiotemp/blogCats.tex}
says ``process only the files you are editing'',
\begin{verbatim}
\input{inputs/inclOnlyCats}
\end{verbatim}
you uncomment a single line in that file to "process  only the files you
are editing".
% Don't comment out files called further down by the
% masterfile \emph{blogCats.tex}, that can get confusing for other people.

\item[2021-12-07 Predrag to Xuanqi].

%You refer to an equation like this: CL18 eq.~{tempBern};

%to figure like this: CL18 figure~{fig:BernCyc2Jacob};

%to table like this: \reftab{tab:LxTs=5/2};

You refer to a reference like this: Gutkin and Osipov\rf{GutOsi15}
(\emph{GutOsi15}
refers to an article listed in \emph{../bibtex/siminos.bib}).

and to external link like this:
``For great wallpapers, see overheads in
\HREF{http://www-personal.umich.edu/~engelmm/lectures/ShortCourseSymmetry.html}
{Engel's} course\rf{Engel11}.''

Pro tip: compile \emph{blogCats.tex} often, as you write, and fix errors as
you write. I had to go all they way back to May to find one of Sidney's
unbalanced ``$\{$'' and make the entire blog compile without errors...

\end{description}

\section{Spring 2021 blog}
\label{sect:xuanqi2021}

\begin{description}

\item[2021-12-08 Predrag]
I believe that the $\phi^4$ is not very different from the $\phi^3$
theory for sufficiently strong stretching parameter. Here are proposed
exercises for you to develop intuition about that:
\begin{enumerate}
  \item
Plot the $3^\cl{}$ Smale horseshoe for $\phi^4$ 1-d forward-in time map,
in the $(\ssp_{\zeit},\ssp_{\zeit+1})$ plane, paralleling
\toChaosBook{figure.caption.292} {fig.~15.5} for $\phi^3$, \ie,
\HenonMap\ \refeq{LC21BWcubic}, \refeq{henlattBWcubic},
\refeq{BWpotential}.

The intuition is topological; 1-d parabola repeller (with parabola height
larger than 1) has the same kind of Cantor set as the complete Smale
horseshoe repeller for the {\HenonMap}. Similarly, I expect
\toChaosBook{figure.caption.191} {fig.~11.4}, here \refeq{symmBimod}, to
capture the topology of the $\phi^4$ repeller.

  \item
For the \HenonMap\ stretching parameter values $a$ larger than
\refeq{COM:SterlHen}, the `critical' value $a_h=5.69931\cdots$ guarantee
a complete horseshoe. What is (very roughly) a corresponding value for the
$\phi^4$ theory?
  \item (harder, not essential as yet)
Quotient the $\Dn{1}$ symmetry for $\phi^4$, as in
\toChaosBook{figure.caption.192} {fig.~11.5}.
\end{enumerate}
Ibrahim and you should form a study group tu understand this -
it's absolutely essential.

%However, I am hoping he will focus on implementing variational methods
%for computing {\lattstate}s.

\item[2021-12-08 Predrag]
The above approaches should be safe for multimodal maps with complete
repelling sets,
and it should work for finite-grammar Smale horseshoe repellers.
Smale's
original horseshoe\rf{smale}, his fig.~1 was unimodal, but he also
explicitly gives our $\phi^4$ bimodal repeller, his fig.~5.

\item[2021-12-07 Predrag]
Parallel reading, while you work on the above:
\refchap{chap:LC21}~{\em LC21: Time reversal for lattice field theories}.

Write your study notes up here.

\item[2021-09-12 to 2021-12-22 Predrag]
I have added my guess \refeq{PCphi4} for the infinite coupling $g$
anti-integrable limit of $\phi^4$ theory. That gives a 3-letter
alphabet $\A=\{-1,0,1\}$. One can use it to find by
continuation any {\lattstate}, at $g$ as low as possible.
`Generalized {\HenonMap}s' AKA $\phi^4$ field theory posts are in
\refsect{sect:phi4latt}~{\em Classical {$\phi^4$} lattice field theory}.

\item[2021-12-?? Xuanqi]











\end{description}

\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\ssp}{\ensuremath{x}}             % lattice site field
\renewcommand{\Ssym}[1]{{\ensuremath{s_{#1}}}}    % Boris
\renewcommand{\Refl}{\ensuremath{\sigma}}             % in DasBuch
\renewcommand{\shift}{\ensuremath{d}}                 % in DasBuch

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\printbibliography[heading=subbibintoc,title={References}]

\ChapterEnd % formatted for ChaosBook.org
