% siminos/spatiotemp/chapter/catMapLatt.tex
% $Author: predrag $ $Date: 2021-12-08 00:39:15 -0500 (Wed, 08 Dec 2021) $

% called by siminos/spatiotemp/blogCats.tex

\Chapter{catMapLatt}{5jan2021}{{\catLatt}}
% \label{c-catMapLatt}  % formatted for ChaosBook.org
\renewcommand{\ssp}{x}

\begin{description}
\item[2016-09-09 Predrag]
I have added this chapter with intention to include it as several examples
in ChaosBook.org.

\item[2020-12-16 Predrag] The abstract of my online
Mathematical Physics Webinar,
Rutgers University - the most attended of the series :)

        \item[\videoLink{YouTube.com/embed/%
9r7wJroEVSA}]  {\em % uploaded              2020-09-22
Spatiotemporal cat - a chaotic field theory
         }
(55 min seminar)

     When I refer to a physical phenomenon -such as motions of a Navier-Stokes fluid- as `chaotic', or `turbulent', I am often told: We understand `chaos' for a system such as Lorenz attractor, but what is a `chaotic' field, a field with infinitely many degrees of freedom?

     The goal of the seminar is to answer this question pedagogically, as a sequence of pencil and paper calculations. First I will explain what is 'deterministic chaos' by walking you through its simplest example, the coin toss or Bernoulli map, but reformulated as problem of enumerating admissible global solutions on an integer-time lattice. Then I will do the same with the 'kicked rotor', the simplest mechanical system that is chaotic. Finally, I will take an infinity of `rotors' coupled together on a spatial lattice to explain what `chaos' or `turbulence' looks like in the spacetime.

      What emerges is a spacetime which is very much like a big spring mattress that obeys the familiar harmonic oscillator field theory equations, the discrete Helmholtz equation (or the tight-binding model), but instead of being `springy', this metamaterial is a dicretization of the Euclidean Klein-Gordon equation, with an unstable rotor at every lattice site, that gives, rather than pushes back, a theory formulated in terms of Hill determinants and zeta functions. We call this mother of all chaotic field theories the `spatiotemporal cat'.

      This is the simplest example of reformulating a space and time translationally invariant, exponentially unstable `turbulent' field theory as a (D+1)-dimensional spatiotemporal system which treats space and time on equal footing. Here there is no `evolution in time': there is only the enumeration of the repertoire of admissible tilings of spacetime by invariant (D+1)-dimensional tori, or `periodic orbits', very much as the partition function of the Ising model is a weighted sum formed by enumerating its lattice states. But that is a story for another seminar.
\\

 And if you don't know,
\HREF{https://www.youtube.com/watch?v=_JZom_gVfuw} {now you know}


\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{chapter/CML}
\input{chapter/Green2d}

\newpage %TEMP
\input{chapter/Lagrangian} %PC 2019-09-05: removed from CL18 article

\newpage %TEMP
\input{chapter/QFTlatt}

\newpage %TEMP
\input{chapter/integLatt}

\newpage %TEMP
\input{chapter/resistors}

\newpage %TEMP
\input{chapter/twots2D}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Elastodynamic equilibria of 2D solids}
\label{sect:2Dsolids}
% Predrag
% extracted to here from siminos/spatiotemp/catMapLatt.tex

{\bf Predrag 2018-08-23} this section is now in {\em book/chapters/mattress.tex},
 removed from here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage %TEMP
\input{chapter/lattLit}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\printbibliography[heading=subbibintoc,title={References}]

\ChapterEnd % formatted for ChaosBook.org
