% siminos/spatiotemp/chapter/Henon.tex
% $Author: predrag $ $Date: 2021-12-24 01:25:20 -0500 (Fri, 24 Dec 2021) $

\Chapter{Henon}{28apr2021}{{\Henlatt}}
% \label{c-Henon}  % formatted for ChaosBook.org


\section{\Henon\ blog}
\label{sect:HenonBlog}

\begin{description}

    \item[2021-02-17 Predrag]
Predrag's  key 2005 result is presumably the time-reversal symmetry
induced, temporal lattice cycle by cycle full square
\refeq{EGfactorSincExcl} and \refeq{EGfactorS}.

    \item[2021-02-15 Predrag]
Added \HenonMap\ examples, mostly for Sidney
(to be edited and returned to ChaosBook):

\refexam{exam:HenonMap}~{\em {\HenonMap}}

% \refexam{exam:cycStabMap}~{\em Stability of cycles for maps}

\refexam{exam:HamHenonMap}~{\em {\Henlatt}}

\refexam{exam:HamHenonJacob}~{\em {\Henlatt} stability}

\refexam{exam:RevHenonMap}~{\em Hamiltonian {\HenonMap}, reversibility}

\refexam{exam:StandMapSymmLin}~{\em Symmetry lines of the standard map}

\refexam{exam:CatMapSymmLin}~{\em Symmetry lines of the cat map}


\end{description}

\subsection{{\Henlatt}; anti-integrable limit}
\label{sect:HamHenonMap}

\begin{description}
  \item[2021-12-23 Predrag]
See also
\refsect{sect:phi4latt}~{\em Classical {$\phi^4$} lattice field theory}

\item[2021-12-22] % Predrag]
\HREF{https://amath.colorado.edu/faculty/jdm/research.html} {Jim Meiss}:
"The concept of anti-integrability was introduced by Aubry and
Abramovici\rf{AuAb90} in 1983 for the standard map (PC: 1983? maybe he
means Aubry and Le Daeron\rf{AuDae83}?), viewed as a linear chain of
particles connected by springs in a periodic potential. They reasoned
that the integrable limit corresponded to vanishing potential energy, so
that the springs dominated giving equal spacing at equilibrium. By
contrast, anti-integrability corresponds to vanishing kinetic energy, so
that particles sit at critical points of the potential. What is most
interesting about this limit is that it is relatively easy, using a
contraction mapping style argument, to show that AI states persist, and
this gives conjugacy to a shift on a symbolic dynamics."

\item[2021-06-04 Predrag]
David Meiss' student David G. Sterling\rf{SterlingThesis99}
much (undeservedly)
un-cited \HREF{https://www.proquest.com/docview/304508605} {PhD thesis},
Univ. of Colorado,
{\em Anti-Integrable Continuation and the Destruction of Chaos} has much
to teach us. He studies \emph{coupled {\HenonMap} lattices} in both
Hamiltonian and Lagrangian formulations; his definition seems pretty much
consistent with our \refeq{SVWhenSTlatt}, though he has a coupling
parameter $c$ used to make spatial couplings weak. The
``anti-integrable'' refers to our choice $a\geq6$, I believe - parameter
regimes in which all of the horseshoe orbits exists.

``Specifying the anti-integrable state for an orbit of a coupled map
lattice requires a multidimensional symbolic object which we call a
symbol tensor.''

His Figures 6.7, 6.18 are reminiscent of my pruning front.

Thesis abstract:
[...] Recurrent phenomena, the
simplest of which is periodic motion, are particularly interesting and
practical objects of study. Scientists have long been captivated by the
near periodic motions of the planets. Among them, \Poincare\ was the
first to truly recognize the importance of periodic solutions in
understanding more complex dynamical behavior. [...] This research has
two complementary aspects. Not only do we develop a technique for
locating periodic orbits in discrete dynamical systems, but we then use
these orbits to study bifurcations, most significantly the global
bifurcation that signals the destruction of chaos. Our technique embodies
the following basic principles:
\\
(1) periodic orbits are conveniently described by a variational principle,
\\
(2) in a special case, the variation principle simplifies, and
\\
(3) continuation from this limiting
case is an effective method for studying periodic orbits. We illustrate
this approach on the \HenonMap.

\item[2021-09-12 Sidney, Predrag]
In the $a\rightarrow\infty$, \emph{anti-integrable} limit\rf{StMeiss98}
the \Henon's original map
\refeq{PCe_ar_pres} goes to $a(\ssp^*)^2=1$,
so
\beq
\ssp_\zeit=\Ssym{\zeit}\ssp^*
    \,,\qquad
\ssp^*=a^{-1/2}
    \,,\quad
\Ssym{\zeit}\in\{-,+\}
\,.
\ee{hen-a-infty}
The small perturbation parameter for the problem is
$\ssp^*=a^{-1/2}$, so replace
\beq
\ssp_\zeit = \Ssym{\zeit}\ssp^* +\hat{\ssp}_\zeit
\,,
\ee{hen-a-infty1}
study the \henlatt\ equations for $\hat{\ssp}_\zeit$.

\item[2021-09-12 Sidney]
The process for perturbations that you're describing sounds very
reminiscent of what is in chapter 11 of Townsend, would that be worth
trying to copy here?

\item[2021-12-22 Predrag]
Moved `generalized {\HenonMap}s' AKA $\phi^4$ field theory posts to
\refsect{sect:phi4latt}~{\em Classical {$\phi^4$} lattice field theory}.

\item[2021-06-04 Predrag]
Read also

Aubry and Le Daeron\rf{AuDae83}
{\em The discrete {Frenkel-Kontorova} model and its extensions.
{I}. {Exact} results for the ground-states} (1983)

Sterling and Meiss\rf{StMeiss98}
{\em Computing periodic orbits using the anti-integrable limit}  (1988)

Aubry and Abramovici\rf{AuAb90},
{\em Chaotic trajectories in the standard map.
{The} concept of anti-integrability},
(1990)

Aubry\rf{aub95ant}
  {\em Anti-integrability in dynamical and variational problems} (1995)

\item[2021-06-04 Predrag]
They are probably deep and good, but I find

Bolotin and MacKay\rf{BolMac97}
{\em Multibump orbits near the anti-integrable limit for {Lagrangian} systems},
(1997)

Bolotin and Treschev\rf{BolTre15}
{\em The anti-integrable limit} (2015)

hard to read. Gave up.


\item[2021-06-04 Predrag]
We all might find the Sect.~III of {Wen's 2014 project}
\wwwcb{/projects/Wen14.pdf} interesting.
\\
Wen says that the original \Henon's\rf{henon} Hamiltonian {\HenonMap} is
equivalent to the harmonic oscillator system. By that he means it can be
interpreted as a kicked driven harmonic oscillator with a nonlinear,
cubic potential kicking term\rf{Heagy92}. If we write anything about \henlatt, we
should include this as a physical motivation.

Check out also Zalmond C. Barney
\HREF{https://fau.digital.flvc.org/islandora/object/fau\%3A3758/datastream/OBJ/view/Derivation_of_planar_diffeomorphisms_from_Hamiltonians_with_a_kick.pdf}
{Master's Thesis}
{\em Derivation of planar diffeomorphisms from Hamiltonians with a kick}.

Butusov \etal\rf{BKPPB18}
{\em Discrete chaotic maps obtained by symmetric integration}
might be of interest for adding a reflection symmetry to the {\HenonMap}.

\item[2021-09-07 Predrag 2 Sidney].

\begin{enumerate}
  \item
Plot lattice field values for $a>>1$, then try to look at a perturbation
theory treatment of this.
  \item
Learn about the
perturbative treatments of field theories.
\end{enumerate}

\item[2021-09-09 Predrag]
Re. 1. above:
I quickly tried to sketch how $a>>1$, did not get anything sensible.
Might be yet another crazy idea that met instant death, do not worry
about it for now.

\item[2021-09-12 Sidney]
% It does work for me now, I'm very sorry about the baby...I'll be more careful in the future.

I did some preliminary calculations of $a>>1$ using the Gallas scaling
(so not quite field theory, but I good test) and as $a$ got larger, the
less variation in lattice site values occurred. There were still
negative, and positive values, but in the large $a$ limit the lattice
site values all approached an equal absolute value. I am not sure if that
is useful, perhaps if we looked at the asymptotic behavior of both large
and small $a$ we could come up with some interesting "asymptotic
\henlatt" theory.

\item[2021-09-12 Predrag]
That sounds very good: in this limit the fields are apparently
$\ssp_\zeit=\Ssym{\zeit}\ssp^*$,
$\Ssym{\zeit}\in\{-,+\}$,
where $-\to0$, $+\to1$ is the binary label corresponding to
lattice site $\zeit$.

Basically you get a theory even simpler than the \templatt.

Perturbation theory for large but finite $a$ would come from replacement
$\ssp_\zeit \to \ssp^*+\epsilon\hat{\ssp}_\zeit$ in the equations,
and ordering terms by powers $\epsilon^k$ and $a^{-\ell}$.
Probably  $a^{-\ell}$ only.

\item[2021-09-12 Predrag]
Do you have the analytic value of $\ssp^*$?

\item[2021-09-13 Sidney]
I have changed my code so that it can be easily switched between
the \Henon\rf{henon} \refeq{PCe_ar_pres}, and Endler and Gallas\rf{EG05}
rescaled \refeq{EG05:ar_pres}, I have added this updated code
\emph{Relaxation Method Henon with Orbit Jacobian.py} to
\emph{siminos/williams/python/relax}.

So, let's see if I understand, if we plug \eqref{hen-a-infty} into the
\Henon\ form \refeq{PCe_ar_pres}, expand and keep only linear terms of
$\hat{\ssp}_\zeit$, we get a new \henlatt\ of form:
\beq\label{pert-henlatt}
\hat{\ssp}_{t+1}+2\Ssym{\zeit}a^{1/2}\hat{\ssp}_\zeit+\hat{\ssp}_{t-1}
    =  %\PCedit{1}
       -(m_{t+1}+m_{t-1})\ssp^*
\,,
\eeq
where $\Ssym{\zeit}$ is determined by the cycle itinerary. Is that the
form you were thinking of? If this is the correct procedure, I can't see
a way of extending this to a second-order perturbation theory, as that is
just reproducing \henlatt.

\item[2021-09-13 Predrag]
I have not thought through your \refeq{pert-henlatt} yet.
Maybe $\hat{\ssp}_\zeit\to\ssp^*\hat{\ssp}_\zeit$ helps a bit.

You have analytic formulas for fixed points \refeq{COM:eq2}, period-2
{\lattstate}s \refeq{COM:2-cycle}. You might find useful approximate
large $a$ formulas for all period-\cl{}\ {\lattstate}s.
They might already be in David Sterling's %\rf{Sterling99}
\HREF{https://www.proquest.com/docview/304508605} {PhD thesis}.

Maybe {\HillDet}s have interesting expansions in powers of
$a^{-\ell/2}$...


\end{description}


For the contraction parameter value
 $b=-1$ the \HenonMap\rf{henon} is orientation and area
preserving, and can be written as a 3-term recurrence
relation
\beq
  x_{\zeit-1} = 1-ax_\zeit^2-x_{\zeit+1}
\,.
\ee{PCe_ar_pres}
Multiply both sides by $-2a$ and define the  \Henon\ `field' at lattice
site $\zeit$ to be $\ssp_\zeit=-2ax_\zeit$. We shall refer to this form
of \Henon\ as the {\em \henlatt}:
\beq
\ssp_{\zeit+1} - \frac{1}{2}\ssp_{\zeit}^2 +\ssp_{\zeit-1}=-2a
\,.
\label{Hen-1dLatt}
\eeq
or we can multiply both sides by $-2$ and define the  \Henon\ `field' at lattice
site $\zeit$ to be $\ssp_\zeit=-2x_\zeit$, yielding {\henlatt} of form:
\beq
\ssp_{\zeit+1} - \frac{a}{2}\ssp_{\zeit}^2 +\ssp_{\zeit-1}=-\Ssym{\zeit}
\,,\quad
\Ssym{\zeit}=2
\,,
\label{Hen-1dLattA}
\eeq
with the `coupling constant' $a$ the analogue of the stretching factor
${s}$ in \templatt\ \refeq{LC21:catMapNewt}, and a constant source
$\Ssym{\zeit}$.

The fixed points $\ssp_j=\ssp$ satisfy
\beq
\ssp^2 - \frac{4}{a}\,\ssp - \frac{4}{a} =0
\,,\qquad
\ssp_\pm =  \frac{2}{a} \pm \sqrt{ \frac{4}{a^2}+  \frac{16}{a}}
\label{Hen-1dLattFix}
\eeq




{\Henlatt} is $\phi^3$ lattice field theory.
Biham-Wenzel\rf{afind} find \Henon\ {\lattstate}s by constructing a cubic
action density \refeq{BWcubic},
\beq
S[\ssp]_\zeit - \Ssym{\zeit}\,\ssp_\zeit
      =  \ssp_{\zeit+1}\ssp_\zeit + \ssp_\zeit\ssp_{\zeit-1}
         -\frac{a}{3!}\ssp_\zeit^3
         - \Ssym{\zeit}\,\ssp_\zeit
\,,\qquad
\Ssym{\zeit}=2
\,.
\ee{henlattBWcubic}

Still to check: is $\Ssym{\zeit}=\pm2$ the Biham-Wenzel method?
Looks like it, as that amounts to flipping the sign of the cubic term
while keeping the variation across there sits of the same magnitude.

I took $+2\ssp_n$ out of $S[\ssp]$ to treat it as a source density term
$\Ssym{\zeit}\ssp_\zeit$.

Now the {\jacobianOrb} \refeq{jMorb1dField} is of the same form as the \templatt\
{\jacobianOrb} \refeq{LC21:Hessian},
\beq
\jMorb[\Xx] =
\left(\begin{array}{cccccccc}
%\begin{pmatrix}
 {s}_{0} &-1 & 0 & 0 & \cdots & 0 & 0 &-1 \\
      -1 & {s}_{1} &-1 & 0 & \cdots & 0 & 0 & 0 \\
       0 &-1 &  {s}_{2} &-1 & \cdots & 0 & 0 & 0 \\
  \vdots & \vdots & \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
       0 & 0 & 0 & 0 & \cdots &-1 &  {s}_{\cl{}-2} &-1 \\
      -1 & 0 & 0 & 0 & \cdots & 0 &-1 &  {s}_{\cl{}-1}
%\end{pmatrix}
          \end{array} \right)
\,,
\ee{jMorb1dField} %was {PCJiKoKr20(8)}
%
but with the stretching factor at site $\zeit$ depending on
the particular \lattstate, ${s}_{\zeit}=\ssp_{\zeit}$,
and once you have an expression for  \HillDet\ $\|\jMorb[\Xx]\|$ in terms of traces
$\Tr\jMorb^k$ ,
\ie, the $\Dn{\cl{}}$ invariant
orbital sums for products of fields on consecutive lattice sites,
they will be the same for the \templatt\ and the \henlatt.

\section{{\HenonMap} symmetries}
\label{s-HenonSymm}
% 2021-02-15  Predrag if edits, return to appendFiniteGr
\index{Henon@H\'enon map!symmetries}
\index{symmetry!H\'enon map}
\index{map!H\'enon}

We note here the symmetries of the {\HenonMap}
\refeq{eq2.1a}.
For $b \neq 0$ the  {\HenonMap} is reversible:
the backward iteration of \refeq{2-step} is given by
\beq
  x_{n-1} = -\frac{1}{b} (1-ax_n^2-x_{n+1})
\,.
\ee{e_henon_inv}
Hence the time reversal amounts to $b \rightarrow 1/b$,
$a \rightarrow a/b^2$ symmetry in the parameter plane, together
with $x \rightarrow -x/b$ in the coordinate plane, and there is
no need to explore the $(a,b)$ parameter plane outside the strip
$b \in \{-1, 1\}$.
For $b=-1$ the map is orientation and area preserving
Hamiltonian {\HenonMap} %\refeq{e_ar_pres}
% restore     \PublicPrivate{,}{% switch \PublicPrivate{
% restore     (see \refeq{} below),
% restore           }% end \PublicPrivate{
% restore     \PC{define orientation, area preserving}
\index{orientation!preserving map}
\index{map!orientation preserving}
\index{area preserving!map}
\index{map!area preserving}
\beq
  x_{n-1} = 1-ax_n^2-x_{n+1}
\,,
\ee{e_ar_pres}
the backward and the forward iteration are the same, and the \nws\ is
symmetric across the $x_{n+1} = x_n$ diagonal. We can write this as a
nonlinear field equation with a Laplacian \refeq{Box(2.1.1)} and a
``cubic'' potential \refeq{BWcubic}
\beq
  % x_{n-1}-2\,x_{n}+x_{n+1}
\Box\,x_{n}+(a\,x_{n}+2)\,x_{n}\,=\, 1
\,.
\ee{HenonLapl}
% restore     \PC{define the \orev\ case}
Endler and Gallas\rf{EG05} prefer the equivalent, rescaled form
\refeq{EG05:ar_pres}.

This is one of the simplest models of a \PoincMap\ for a Hamiltonian flow.
\index{Hamiltonian!flow}
\index{flow!Hamiltonian}

For the \orev\ $b=1$ case we have `golden \Henon' (in analogy with \refeq{goldenCat})
\index{orientation!reversing map}
\index{map!orientation reversing}
\beq
  x_{n-1} = 1-ax_n^2+x_{n+1}
\,,
\ee{e_ar_rev}
and the \nws\ is symmetric across the $x_{n+1}=-x_n$ diagonal.

  \input{Solutions/soluCOM}

\section{Symmetries of the \topp}
\label{c-symbol-plane}
% 2021-02-15  Predrag if edits, return to appendFiniteGr

% restore     \cyclist

    \PC{2021-04-03}{
Moved to  here from ChaosBook \emph{appendFiniteGr}. Return once updated here.
    }
Depending on the type of dynamical system, the \topp\ might have
a variety of symmetries.
Under the time reversal
\beq
   \biinf{\Ssym{-2}\Ssym{-1}\Ssym{0}}{\Ssym{1}\Ssym{2} \Ssym{3}}
\to
   \biinf{\Ssym{3}\Ssym{2}\Ssym{1}}{\Ssym{0}\Ssym{-1} \Ssym{-2}}
\ee{COM:timeRevSymb}
the points in the  \topp\ for an \opres\ map are
symmetric across the diagonal $\gamma=\delta$, and
for the \orev\ case they are symmetric
with respect to the $\gamma=1-\delta$ diagonal.
Consequently the  periodic orbits appear either in dual pairs
$p=\Ssym{1} \Ssym{2} \Ssym{3} \ldots \Ssym{n}$,
$\overline{p}=\Ssym{n} \Ssym{n-1} \Ssym{n-2} \ldots \Ssym{1}$,
or are self-dual under time reversal,
$S_{p} = S_{\overline{p}}$.
% $\Ssym1 \Ssym2 \Ssym3 \ldots \Ssymn = \Ssymn \Ssym{n-1} \Ssym{n-2} \ldots \Ssym1$.
For the \opres\ case
a self-dual cycle of odd period has at least one point on the
symmetry diagonal. In particular, all fixed points lie on the symmetry
diagonal. Determination of such symmetry lines can be of considerable
practical utility, as it reduces some of the periodic orbit
searches to 1-dimensional searches.
	\PC{2021-03-24} {
create appendix
	chapter/appendCont.tex,
include
add JH Jan 18, 2008
{\em Desymmetrization of large spaces,
     thinking is extra price version}
% \label{sect:deSymmThink}
from
	halcrow/blog/TEX/symm.tex;
create
	Problems/exerAppCont.tex,
	Problems/soluAppCont.tex,
include
	halcrow/blog/TEX/zeghlache.tex;
create
	 chapter/refsAppCont.tex.
	}

\subsection{Symmetry lines}
\label{sect:SymmLines}
discuss symmetry lines

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \example{Symmetry lines of the standard map.}
\hfill \fastTrackExam{exam:StandMapSymmLin}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \example{Symmetry lines of the cat map.}
\hfill \fastTrackExam{exam:CatMapSymmLin}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\printbibliography[heading=subbibintoc,title={References}]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage % REMOVE
  \input{../spatiotemp/Examples/examHenonMap}
  \input{../spatiotemp/Examples/examHamHenonMap}
  \input{../spatiotemp/Examples/examHamHenonJacob}
  \input{../spatiotemp/Examples/examRevHenonMap}
  \input{../spatiotemp/Examples/examStandMapSymmLin}
  \input{../spatiotemp/Examples/examCatMapSymmLin}

  \input{Problems/exerHamHenonCyc}
  \input{Solutions/soluHamHenonCyc}

\ChapterEnd % formatted for ChaosBook.org
