% siminos/cats/nonlin-v1/reviews/response.tex
% $Author: predrag $ $Date: 2020-10-09 12:02:16 -0400 (Fri, 09 Oct 2020) $

% Predrag                                       2020-10-09
% Boris                                         2020-09-25

                        %% logical setup, no need to edit %%%%%%%%%%
                        \newif\ifpaper \newif\ifPDF               %%
                        \newif\ifOUP                              %%
                        \newif\ifboyscout\boyscouttrue      %% commented %%
                        \newif\ifdasbuch \newif\iftoCB            %%
                        \newif\ifsolutions \newif\ifblog          %%
                        \newif\ifsubmission\submissionfalse %% internal  %%
                        \blogtrue                                 %%
                        \dasbuchtrue %% DasBuch, not QFT lectures %%
                        \solutionstrue %% include solutions       %%
                        \paperfalse\PDFtrue %% hyperlinked        %%
                        \OUPfalse \toCBtrue      %% ChaosBook %%%%%%
% Toggle between draft and public versions:
% \boyscoutfalse                        % public, hyperlinked
% \boyscoutfalse\submissiontrue         % for Nonlinearity
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\documentclass[12pt]{iopart} %{article}
\usepackage{iopams}
\pdfminorversion=4

\usepackage[pdftex]{color}
%\newenvironment{thebibliography}[1]{}
%\usepackage[numbers]{natbib} % Predrag 2010-06-08: to avoid
% Package natbib Error: Bibliography not compatible with author-year citations
\usepackage{amsmath,amsfonts,amssymb,amsbsy,amscd,amsgen}
\usepackage{url}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex,colorlinks]{hyperref} %% hyperlinks, LAST package called
\input{../../inputs/biblatex}
    \addbibresource{../../bibtex/siminos}
%\input{../../inputs/def}
\input{../defsCats}      %% all article-specific edits: \renewcommand, etc
\graphicspath{{../../Fig/}{../../figs/}}

\newtheorem{rmark}{{\small\textsf{\textbf{Remark}}}}[section]
\newcommand{\remark}[2]{
        % \begin{quotation}
        \begin{rmark}
%        {\small\em\noindent {\small\sf \underline{ #1} ~} #2 }
        {\small\em\noindent {\small\sf #1 ~} #2 } % 2010-12-12 experiment
    \end{rmark}
    % \end{quotation}
              }
   \newcommand{\TWRUN}[2]{\ensuremath{\textrm{TW}_{#1}}}
   \newcommand{\RPORUN}[2]{\ensuremath{\textrm{RPO}_{#1}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagestyle{plain}
\renewcommand{\baselinestretch}{1.0}


\begin{document}
\bigskip
                    \hfill October 09, 2020

\noindent
XXX, Editor-in-Chief\\
Nonlinearity

\bigskip
\bigskip

\noindent
Dear Professor XXX

\bigskip

\BG{2020-09-25}
{First referee is easy case to solve, so I do not comment on this.}

\BG{2020-09-25}
{The second one is a real nut case.
The  detailed list of corrections is not a problem, but he also wants:

1. The meaning of some key constructs is obscured by poor notation and
terminology.
    \\
2. Some key results should be formalized as theorems; some numerical
results should be distilled into conjectures.

These two issues seem to me disconnected. The first one he describes in
details, so it might be adjusted to his satisfaction (whether it makes
sense or not).

The second issue is really annoying, as he does not describes precisely
what he wants.  The guy partitions the world into lemmas and theorems,
otherwise  nothing is comprehensible.
}



The bottom line. There are two things which make our results non
rigorous:

  1)
In  the introductory part we state  that frequencies of block appearances
are given by the measures of the corresponding cylinder sets.  The
argument is based on the ergodic theorem,  but we do not prove  that  our
natural (Lebesgue) measure is uniquely ergodic. We only show that the
system is fully hyperbolic. For physicist it would be enough, but not for
a mathematician since there are (typically crazy) examples of hyperbolic
systems  which are not uniquely ergodic. Our claim on p.4 '' The d = 2
spatiotemporal cat is fully hyperbolic and ergodic for s > 4'' is
actually conjecture (regarding ergodicity). This is in principle
incurable (would require a separate work), however we do have numerical
confirmation of this. The main  value of our numerics is exactly about
this -  we confirm that the time averages (frequencies) converge to the
measures of the sets  which we calculate analytically. Another conjecture
which we can put forward is extension of the results to d>2.

2)
In the body of the paper, where we calculate measures  $\mu(M_\R)$
everything is in principle rigorous. But we do not present it  in a
mathematical clean way for d=2 case. For this we need to define  more
clearly what is our measure $\mu$ and what is  $\mu(M_\R)$. In principle we
can convert then eq. 44 into a  ``theorem/proposition`` and   3
subsections after (interior symbols, rationality, normalization) into
corollary. The same goes (more easily) for single cat.   Another possible
''theorem/proposition '' is eq. line between eq.(40) and eq. (41), resp.
eq. 26 for single cat.

    \BG{2020-09-25}
    {Do you agree with the above synopsis?}
    \PC{2020-10-09}
    {In principle, yes. But I have not read the reports yet}
    \PC{2020-10-09}
    {
 I would like to redefine the stretching factor s from\\
    current:    $(\Box + s - 2d) x_z = m_z$     in this paper, cats/GHJSC16.tex\\
    proposed:   $(\Box + d(s - 2)) x_z = m_z$   used in kittens/CL18.tex\\
    why? Because it is the right thing to do. Because $s=2$ is the
    marginal dimension in all of the literature (Gaussian models, etc.).
    Briefly, the Laplacian, which corresponds to $s=2$, has only a single
    marginal eigenvalue in all dimensions, hence all d-torus periodic
    states counting has a single prefactor od $(s-2)$ in all dimensions.
    }


% \newpage
\printbibliography[
heading=bibintoc,
title={References}
				  ] %, type=online]  % if not using default "Bibliography"

\end{document}
