% siminos/posters/GTretreat07.tex
% $Author: siminos $ $Date: 2008-01-01 18:54:45 -0500 (Tue, 01 Jan 2008) $

\documentclass{article}
\usepackage[usenames]{color}
\pagestyle{empty}

\input{defs}


\begin{document}

{\normalsize \bf \color{NavyBlue} Relative periodic orbits in Kuramoto-Sivashinsky Equation}
\newpage

{\Large \color{green} Evangelos Siminos and Predrag Cvitanovi\'c}
\newpage

\begin{minipage}[t]{7in}
	\begin{center}
		{\normalsize School of Physics, Georgia Institute of Technology,\\
		Atlanta, GA 30332, USA }
	\end{center}
\end{minipage}

\newpage

{\Large \color{green} Ruslan L. Davidchack}
\newpage

\begin{minipage}[t]{7in}
	\begin{center}
		{\normalsize Department of Mathematics, University of Leicester,\\
		University Road, Leicester LE1 7RH, U.K. }
	\end{center}
\end{minipage}

\newpage

\Large
\begin{minipage}[t]{7in}
\begin{center}
The \KS\ partial differential equation (\kse):
\[
	u_t=2uu_x-u_{xxxx}-u_{xx} \,,
	\label{eq:KSe}
\]
$\ x\in[0,L]$ with periodic boundary conditions $u(0,t)=u(L,t)$.
\end{center}
\end{minipage}

\newpage

 {\color{NavyBlue}\LARGE\bf Appears in the description of}

\newpage

\begin{minipage}[t]{7in}
\begin{itemize}
	\item hydrodynamic instabilities in laminar flame fronts,
	\item reaction-diffusion systems,
	\item the interface between two viscous fluids,
	\item the evolution of liquid falling films.
\end{itemize}
\end{minipage}

\newpage
{\color{NavyBlue}\LARGE\bf Our motivation for \kse\ study}

\newpage

\begin{minipage}[t]{7in}
 \begin{center}
 It is one of the simplest PDE's that\\ 
 exhibit spatiotemporal chaos.
 \end{center}
\end{minipage}

\newpage
{\color{NavyBlue}\LARGE\bf Our approach }
\newpage 
{\color{NavyBlue} Use dynamical system language and tools. }
\newpage 
Study
\newpage

\begin{minipage}[t]{6in}
\begin{itemize}
	\item Phase space dynamics, 
	\item Organization of dynamics by invariant objects\\ 
	(equilibria, periodic orbits, their stable/unstable manifolds)
	\item Role of symmetries: 
\end{itemize}
\end{minipage}

\newpage
\begin{center}
Translational\\ invariance  
\end{center}
\newpage
\begin{center}
relative equilibria\\ and\\ relative periodic orbits.
\end{center}
\newpage
Goal
\newpage

\begin{minipage}[t]{7in}
	\begin{center}
		Apply Periodic Orbit Theory to predict \\
		statistical quantities for specific system size.
	\end{center}
\end{minipage}

\newpage

{\color{NavyBlue}\LARGE\bf  Phase space v.s. physical space }

\newpage

\begin{flushleft}
	\begin{tabular}{rl}
		PDE: &  $\infty$-dimensional dynamical system \\
	
		Transform to Fourier space: & system of $\infty$-many, nonlinearly coupled ODE's \\
	
		Dissipation: & only a finite number of modes active \\
	
		Truncate system: & $d$-dimensional system of nonlinearly coupled ODE's
	\end{tabular}
\end{flushleft}

\newpage

\LARGE{\color{NavyBlue}\LARGE\bf KSe for L=22.0}

\newpage

\begin{minipage}[t]{5in}
Unstable Equilibria and Relative Equilibria (traveling waves) for $L=22.0$. 
From left to right: $2$-cell equilibrium, $3$-cell equilibrium, 1-cell traveling
wave (to the left) and it's $u(x,t)\rightarrow -u(-x,t)$ counterpart. The traveling
waves are shown decaying to a typical chaotic state.
\end{minipage}

\newpage

\begin{minipage}[t]{5in}
Unstable manifold of 2-cell equilibrium of KSe
for $L=22.0$. The black line represents the family of 2-cell
equilibria obtained by application of the translation operator, 
while the purple line represents the family of 3-cell equilibria.
The red trajectory represents the heteroclinic connection from a 2-cell equilibrium 
to a 3-cell equilibrium. The connection splits the manifold into two parts, 
colored blue and green here.
\end{minipage}

\newpage

\begin{minipage}[t]{5in}
Relative Periodic Orbits satisfying the condition $u(x+\Delta,t+T)=u(x,t)$ for time period $T$ 
and spatial displacement $\Delta$. %The periodic orbit of $T=47.64$ visit's the neighborhood of
%two members of the 2-cell equilibria family, while the one of period $T=76.60$ visit's the ne
\end{minipage}

\newpage

{\color{NavyBlue}\LARGE\bf  Future }

\newpage

\begin{minipage}[t]{7in}
 \begin{itemize}
  \item Construct symbolic dynamics
  \item Find all periodic and relative periodic orbits up to a given period
  \item Use trace formulas that incorporate continous symmetries
 \end{itemize}
\end{minipage}

\newpage
	\refref{ku}
	\refref{siv}
	\refref{DasBuch}
	\refref{lanCvit06}
	\refref{SCD06}
	\refref{Christiansen:97}
\newpage

\begin{minipage}[t]{7in}
\normalsize
\bibliography{../../bibtex/siminos,../../bibtex/PC}
\end{minipage}


\end{document}
