% siminos/blog/bronsky-2005.tex
% $Author: predrag $ $Date: 2013-11-15 15:54:52 -0500 (Fri, 15 Nov 2013) $

\section{Uncertainty estimates and $L_2$ bounds for the KS}

% Predrag moved from thesis/chapters/ back to blog  oct  2 2009
% Predrag moved from blog/ to thesis/chapters/      jun 26 2008
{\bf
Jared C. Bronski
%\footnote{Department of Mathematics,
% University of Illinois Urbana-Champaign, 1409 W. Green St, Urbana IL 61801}\\
and
Tom Gambill} references, not yet BibTeXed:

M.S. Jolly, I.G. Kevrekidis and E.S. Titi,
Approximate inertial manifolds for the {Kuramoto-Sivashinsky} equation:
analysis and computations,
Physica D 44 (1990), 38-60.

E. Mitidieri and S.I. Pohozaev,
Apriori Estimates and Blow-up of Solutions to Nonlinear Partial
Differential Equations and Inequalities,
Procedings of the Steklov Institute of Mathematics Issue 3 vol 234
(2001).

Y. Pomeau and P. Manneville,
Stability and fluctuations of spatially periodic flow,
J. Physique Lett. 40 (1979), 609-612.

P. Souplet,
Gradient blow-up for multidimensional nonlinear parabolic equations with
general boundary conditions,
Di. Inegral Eqns 15 (2002), 237-56.

E. Tadmor, The well-posedness of the {Kuramoto-Sivashinsky} equation,
SIAM Journal on Mathematical analysis 17 (1986), 884-893.

P. Constantin, C. Foias, B.  and R. Temam\rf{constantin_integral_1989}.

article{Foias}
Foias, C., Manley, O. and Temam, R. (1987).
Attractors for the {B\'{e}nard} problem:
{Existence} and physical bounds on their fractal dimensions.
{\it Nonlinear Anal.}
{\bf 11}, ~939-967.

article{Foias2}
Foias, C., Sell, G. R., and  Temam, R. (1988).
Inertial manifolds for nonlinear evolutionary equations.
{\it J. Diff. Eq.}
{\bf 73} (2), ~309-353.

C. Foias and I. Kukavica, Determing nodes for the {Kuramoto-Sivashinsky} equation,
J. Dynam. Di. Eq. 7 (1995), 365-373.



article{LMRT} LaQuey, R., Mahajan, S., Rutherford, P. and Tang, W. (1975).
{\it Phys. Rev. Lett} {\bf 34} (7) 391-394.

article{ST} Sell, G. and Taboada, M. (1992).
 Local dissipativity and attractors for the {Kuramoto-Sivashinsky}
equation in thin $2{\rm D}$ domains.
{\it Nonlinear Analysis} {\bf 18} (7), 671-687.
