% siminos/cats/nonlin-v1/reviews/response2-v1.tex
% $Author: predrag $ $Date: 2020-11-17 00:57:43 -0500 (Tue, 17 Nov 2020) $

% Predrag                                       2020-11-16
% Predrag                                       2020-10-26
% Boris                                         2020-09-25

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
                        \newif\ifboyscout\boyscouttrue      %% commented %%
                        \newif\ifsubmission\submissionfalse
                        \newif\ifhighlightedits\highlighteditstrue
% Toggle between draft and public versions:
% \boyscoutfalse\highlighteditsfalse    % public, hyperlinked
% \boyscoutfalse\highlighteditstrue     % for Nonlinearity referees
% \boyscoutfalse\submissiontrue         % for Nonlinearity
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{iopart}
% loads AMS amsgen, amsfonts, amsbsy, amssymb:
\usepackage{iopams} % to load AMS extension fonts msam and msbm
                     % the blackboard bold alphabet, extra maths symbols
                     % and extra definitions for bold Greek letters.
                     % do not use amsmath.sty

\pdfminorversion=4  % the very start the TeX file, so PDF or bitmap figures
                    % are PDF version 1.4 or lower

\usepackage{graphicx}              % is the recommended load line
\graphicspath{{../figs/}{../Fig/}}  %% directories with graphics files
\usepackage{color} % dvips allows for colors
\usepackage[colorlinks]{hyperref} %% hyperlinks
\usepackage{amsthm}
\usepackage{parskip}
\input{../biblatex}    % this makes references hyperlinked
\addbibresource{../../bibtex/siminos.bib}

\input{../defsCats}      %% all article-specific edits: \renewcommand, etc

\pagestyle{plain}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{2em}

\begin{document}
                    \hfill October 25, 2020

\bigskip
\noindent
---------------------------------------------------------------------- \\
Referee 2 Report -- NON-104197 / Gutkin et al \\
---------------------------------------------------------------------- \\

\bigskip
\bigskip



\noindent
In what follows, referee's remarks are followed by our responses, either
as indented ``{\em Response~~~} $\cdots$'' blocks of text, or marked
\edit{blue}. In the attached revised manuscript, all edits
are indicated by \edit{blue text}.
\bigskip\bigskip

The referee writes:
I would like to add a few questions that came to my mind and might
be still considered in the manuscript:

\begin{enumerate}
  \item[(1)]
bottom of p. 5. Is it clear why the system is chaotic for
    $s > 2$ ?
    What is the reason for such a simple relation with the
    dimensionality?

\begin{quote}
{\em Response~~~}
To answer this key question clearly,
we have now redefined the stretching factor $s$ from the initial
Nonlinearity submission\\
  $(\Box + s - 2d) x_z = m_z$  to
\beq
\edit{(\Box + d(s - 2)) x_z = m_z}
\ee{kittensEq}
Why?

We do not believe it would be wise to get into the larger picture in this
paper (whose focus is narrow - evidence for the conjectured \catlatt\
ergodicity), so the remarks that follow are just for the referee:

\catLatt's continuum relatives are
the inhomogeneous \emph{Helmoltz equation}\rf{DiHaHu01,Lick89,FetWal03}
\beq
   (\Box+k^2)\,\field(x)= -4\pi\rho(x)\,,\qquad x\in \reals^d
\,,
\label{CatMapContinuesPC}
\eeq
and, for
the ${\mu}^2=-k^2>0$ (imaginary $k$) case, the equation
\beq
   (\Box-{\mu}^2)\,\field(x)= -\rho(x)\,,\qquad x\in \reals^d
\,,
\label{sPe}
\eeq
known as  the
{\sPe}, Klein–Gordon or Yukawa equation\rf{Dorr70,GoVanLo96,HuCon96,HuRyCo98}.
Sitting between the two is
 the Poisson equation, the $k \to 0$ limit of the Helmholtz equation;
\beq
   \Box\,\field(x)= -4\pi\rho(x)\,,\qquad x\in \reals^d
\,.
\ee{PoissonEq}
For $\rho=0$, the equation is known as \emph{Laplace's equation}.

The $d$\dmn, purely hyperbolic ${\mu}^2>0$
{\catlatt} \refeq{kittensEq}
\beq
 (\Box - {\mu}^2\unit)_{zz'} \field_{z'} = \m_z
    \,, \qquad
  \field_{z} \in  \mathbb{T}^{1}
    \,, \quad
  m_{z} \in \A^{1}
    \,, \quad
  z\in \integers^{d}
\,,
\ee{GreenLinearConnPC}
that we study is a discretization of the inhomogeneous {\sPe}
\refeq{sPe},
where the Yukawa massive field mass parameter is related to the \catlatt\
\refeq{kittensEq} stretching parameter ${s}$ by
\beq
{\mu}^2=d(s-2)
\,.
\ee{catlattMass}
In this parametrization it is clear
that \catlatt\ is hyperbolic for $s>2$ in any dimension $d$.
\end{quote}

  \item[(2.a)]
Do the results for classical discrete cat-lattices allow implications for
the continuum limit, ie. where the system passes from a lattice to a
field?

\begin{quote}
{\em Response~~~}
For $|s|<2$ the limit is the Helmholtz equation
\refeq{CatMapContinuesPC}, with oscillatory eigenmodes.
The difference between the three cases is illustrated by
the $d=3$ dimensions' outgoing Green's functions for $|s|<2$,
$s=2$ and $s>2$, respectively:
\bea
g_+({x},x') &=& -\frac{e^{ik\vert{x}- x'\vert}}{4\pi\vert{x}- x'\vert}
\continue
g({x},x') &=&-\frac{1}{4\pi\vert{x}- x'\vert}
\continue
g({x},x') &=& -\frac{e^{-{\mu}\vert{x}- x'\vert}}{4\pi\vert{x}- x'\vert}
\,.
\label{GreenFunContinuesPC1}
\eea
\end{quote}

  \item[(2.b)]
What is expected for the coding in the continuum limit?

\begin{quote}
{\em Response~~~}
The $s>2$ continuum limit of the coding, or the field $\field(x)$ we
have not thought through carefully.
The quick answer
is {\sPe} \refeq{sPe} with exponentially decaying eigenmodes, but the
{\catlatt} \refeq{kittensEq} $\mod 1$ restriction $\field_{z} \in
\mathbb{T}^{1}$ (rather than the usual {\sPe} scalar field $\field_{z}\in\reals$)
might require some care.
\end{quote}

  \item[(3.a)]
Do we need to consider $s\to\infty$ to keep chaotic dynamics?

\begin{quote}
{\em Response~~~}
No, the system is chaotic for any $s>2$, but with a
            nontrivial grammar for admissible symbol \brick s.
\end{quote}

  \item[(3.b)]
Does this imply simple Bernoulli processes in the $s\to\infty$  limit?

\begin{quote}
{\em Response~~~}
     Yes, for $s\to\infty$, the ``inner", ``bulk" full-shift alphabet
     overwhelms the ``outer", ``boundary", and the system is simple
     Bernoulli process in the same sense that the usual $1d$ Bernoulli
     map generates a simple Bernoulli process.
\end{quote}

  \item[(4)]
May it happen that one is always running into a not fully chaotic regime
when fixing the local dynamics parameter $s$ and increasing the
dimension?

\begin{quote}
{\em Response~~~}
    No, with $s$ defined in \refeq{kittensEq}, the {\catlatt} is chaotic
    for $s>2$ in all $d$.
\end{quote}

  \item[(5)]
 What are implications for semiclassical many-body approaches?

\begin{quote}
{\em Response~~~}
As explained in Gutkin and Osipov\rf{GutOsi15}, the partner
periodic tori considered in Sect.~5.2 play a major role in the
semiclassical treatment of the  quantum spatiotemporal  cat problem since
their action differences are small.    The results of the current paper
allow for an effective way to construct them. Their  qualitative
contribution into spectral correlations  is a challenging  problem which
is beyond the scope of the current submission.
\end{quote}
\end{enumerate}

% \newpage
\printbibliography[
heading=bibintoc,
title={References}
				  ] %, type=online]  % if not using default "Bibliography"

\end{document}
