% siminos/cats/nonlin-v2/response-v1.tex
% $Author: predrag $ $Date: 2020-11-17 00:57:43 -0500 (Tue, 17 Nov 2020) $

% Predrag                                       2020-11-16
% Predrag                                       2020-10-26
% Boris                                         2020-09-25

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
                        \newif\ifboyscout\boyscouttrue      %% commented %%
                        \newif\ifsubmission\submissionfalse
                        \newif\ifhighlightedits\highlighteditstrue
% Toggle between draft and public versions:
% \boyscoutfalse\highlighteditsfalse    % public, hyperlinked
  \boyscoutfalse\highlighteditstrue     % for Nonlinearity referees
% \boyscoutfalse\submissiontrue         % for Nonlinearity
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{iopart}
% loads AMS amsgen, amsfonts, amsbsy, amssymb:
\usepackage{iopams} % to load AMS extension fonts msam and msbm
                     % the blackboard bold alphabet, extra maths symbols
                     % and extra definitions for bold Greek letters.
                     % do not use amsmath.sty

\pdfminorversion=4  % the very start the TeX file, so PDF or bitmap figures
                    % are PDF version 1.4 or lower

\usepackage{graphicx}              % is the recommended load line
%\usepackage[pdftex]{graphicx}  % PC 2019-12-05 arXiv error
%\usepackage[pdflatex]{graphicx} % PC 2019-12-05 suggested by arXiv error
\graphicspath{{../figs/}{../Fig/}}  %% directories with graphics files

    \ifsubmission
% for submission, copy figures into /nonlin-v2/, then
% (not required) comment out the \graphicspath{
    \else
% prepare hyperlinked
\usepackage{color} % dvips allows for colors
\usepackage[colorlinks]{hyperref} %% hyperlinks
\usepackage{amsthm}
\input{../biblatex}    % this makes references hyperlinked
\addbibresource{../../bibtex/siminos.bib}
     \fi

\input{../defsCats}      %% all article-specific edits: \renewcommand, etc



\pagestyle{plain}
\renewcommand{\baselinestretch}{1.0}


\begin{document}
\bigskip
                    \hfill November 15, 2020


\bigskip
\noindent
---------------------------------------------------------------------- \\
Revision 1 -- NON-104197 / Gutkin et al \\
---------------------------------------------------------------------- \\
\bigskip
\bigskip

\noindent
Dear editor

attached is the revision 1 of our paper that incorporates the edits suggested
by the referees, together with the our responses to the referee reports.

Just one kvetch, of principle. Referee 1 is of the opinion that

\begin{quote}
{\em $\cdots$ the authors should attempt to convert the main results into
theorems. If gaps of rigour obstruct this process, these should be
spelled out clearly.}
\end{quote}


\emph{Nonlinearity} is a joint project of a math and a physics society
for a reason. It is not a people's reeducation camp whose purpose is to
turn physicists into mathematicians. We are here to meet here as sisters,
loving, curious and respectful of our respective cultures and
backgrounds. It would never occurre to me to demand of a mathematician
the she should write in the style customary to - lets say physical
chemists.
To cite our Constitution, so thoughtfully phrased by our eponymous fathers:

\emph{Nonlinearity} scope: Aimed primarily at mathematicians and physicists
interested in research on nonlinear phenomena, the journal's coverage
ranges from proofs of important theorems to papers presenting ideas,
conjectures and numerical or physical experiments of significant physical
and mathematical interest.

We have now reformated the single cat statements around eqs.~(27) to~(31)
into Theorem~3.1, and the spatiotemporal cat statements around eqs.~(42)
to~(45) into Theorem~4.1, to humor the  referee.
But honestly, we doubt that such faux ``theoremification'' serves anybody
well.
\bigskip\bigskip

\noindent
sincerely
\\
Predrag Cvitanovi\'c

\end{document}
