% intro.tex
% $Author: siminos $ $Date: 2007-10-16 12:02:45 -0400 (Tue, 16 Oct 2007) $

\section{Introduction}

Recent experimental and theoretical advances\rf{science04}
support a dynamical vision of turbulence:
For any finite  spatial resolution,
a turbulent flow follows approximately for a finite time
a pattern belonging to a
{ finite alphabet}
of admissible patterns.
The long term dynamics is
a {walk through the space of these unstable patterns}.
The question is how to characterize and classify such patterns?
Here we follow the seminal Hopf paper\rf{hopf48}, and  visualize
turbulence not as  a sequence of
spatial snapshots in turbulent evolution,
but as a trajectory in an
 $\infty$-$d$ \statesp\ in which an
instant in turbulent evolution is
a {unique} point. In the dynamical systems approach,
theory of turbulence for a given system, with given boundary conditions,
is given by the
(a) geometry of the \statesp\ and (b) the associated natural measure,
\ie,
the likelihood that asymptotic dynamics visits a given \statesp\ region.

We pursue this program in context of the \KS\ (KS) equation,
one of the simplest physically interesting spatially extended
nonlinear systems.  Holmes, Lumley and Berkooz\rf{Holmes96} offer a
delightful discussion of why this system deserves study as a staging
ground for studying turbulence in full-fledged Navier-Stokes
boundary shear flows.

Dynamical \statesp\ representation of a PDE is $\infty$-dimensional,
but the KS flow is strongly contracting and its non-wondering set,
and, within it, the set of invariant solutions investigated here, is
embedded into a finite-dimensional inertial manifold\rf{FNSTks85} in
a non-trivial, nonlinear way. `Geometry' in the title of this paper
refers to our attempt to systematically triangulate this set in
terms of dynamically invariant solutions (\eqva, \po s, $\ldots$)
and their unstable manifolds, in a PDE representation and numerical simulation algorithm 
independent way. The goal is to describe a given
`turbulent' flow quantitatively, not model it qualitatively by a
low-dimensional model. For the case investigated here, the \statesp\
representation dimension $d \sim 10^2$ is set by requiring that the
exact invariant solutions that we compute are accurate to $\sim
10^{-5}$.
The \statesp\ is high-dimensional, the asymptotic dynamics is
confined to a low-dimensional subspace, so we are in the gray,
ill-defined overlapping zone between `turbulence'
and `spatiotemporal chaos,' the two terms that we shall use
interchangeably in what follows.

In previous work, the \statesp\ geometry and the natural measure for
this system have been
studied\rf{Christiansen97,LanThesis,LanCvi07} in terms of unstable
periodic solutions restricted to the antisymmetric subspace of the
KS dynamics.

The focus in this paper is on the role continuous symmetries play in
spatiotemporal dynamics.
The notion of exact periodicity in time is
replaced by the notion of relative spatiotemporal
periodicity, and \reqva\ and \rpo s here play the role the \eqva\
and \po s played in the earlier studies. Our search for \rpo s in KS system was inspired by Vanessa
L{\'o}pez\rf{lop05rel} investigation of {\rpo s} of the Complex
Ginzburg-Landau equation.  However, there is a vast literature on
{\rpo s} since their first appearance, in Poincar\'e study of the
3-body problem\rf{ChencinerLink,rtb}, where the Lagrange points are
the \reqva.  They arise in dynamics of systems with continuous
symmetries, such as motions of rigid bodies, gravitational $N$-body
problems, molecules and nonlinear waves.
Very recently Viswanath\rf{Visw07b}
has found both \reqva\ and \rpo s in the plane Couette problem.

Building upon the pioneering work of \refrefs{KNSks90,ksgreene88},
we undertake here
a study of the \KS\ dynamics
for a specific system size $L = 22$, sufficiently large
to exhibit many of the features typical of `turbulent' dynamics
observed in large KS systems, but small enough to lend itself to a
detailed exploration of the  \eqva\ and \reqva,
their stable/unstable manifolds,
determination of a large number of
\rpo s, and a preliminary exploration of the relation between the
observed spatiotemporal `turbulent' patterns and the \rpo s.

In presence of a continuous symmetry any solution belongs to a group
manifold of equivalent solutions. The problem: If one is to
generalize the periodic orbit theory to this setting, one needs to
understand what is meant by solutions being nearby (shadowing) when
each solution belongs to a manifold of equivalent solutions. We
resolve here this puzzle by demonstrating that if one picks any
particular solution, the universe of all other solutions is rigidly
fixed through a web of heteroclinic connections between them. This
insight garnered from study of a 1-dimensional \KS\ PDE is more
remarkable still when applied to the plane Couette flow\rf{GHCW07},
with 3-$d$ velocity fields and two translational symmetries.


The main results presented here are: (a) Dynamics visualized through
physical, symmetry invariant observables, such as `energy,'
dissipation rate, \etc,
and through
projections onto dynamically invariant, PDE-discretization
independent \statesp\ coordinate frames, \refsect{sec:energy}. (b)
Existence of a rigid `cage' built by heteroclinic connections
between \eqva, \refsect{sec:L22}. (c) Preponderance of
unstable \rpo s and their likely role as the skeleton underpinning
spatiotemporal turbulence in systems with continuous symmetries,
\refsect{sec:rpos}.
