% acknowledgements.t ex
% $Author: predrag $ $Date: 2009-04-06 03:51:05 -0400 (Mon, 06 Apr 2009) $

I would like to thank my advisor, Predrag Cvitanovi\'c,
for his constant encouragement and support, for the
time he put in my education and advisement, and for his patience with
some of my bolder ideas.
Numerics did rarely converge until Ruslan L. Davidchack stepped in
and taught me how to do things correctly, always providing
useful advice and totally changing the route of this thesis.
Yueheng Lan has put a lot of effort in bequeathing his knowledge
and experience in the problem, and I am grateful to him for pointing out to the
existence of the \EQV{1}~\eqv\ at the $L=22$ system size among many
other tips. I have
profited from John F.~Gibson's insights into \statesp\
visualization of high-dimensional fluid flows,
discussions with
Jonathan Halcrow on desymmetrization of Lorenz and other
flows, and
I thank
G.~Robinson,~Jr.\,, National Science Foundation grant
DMS-0807574 and Gerondelis Foundation
for partial financial support. I thank Loukas Vlahos for
initiating my interest in nonlinear dynamics during my
undergraduate studies and his invaluable help in starting this journey.

I would like to thank my parents for their firm support and love that
laid down the foundations for everything. Kalli has been cheerful company while I was
writing this thesis and her keystrokes helped randomize the content.
%
% and now an ode to Penelope:
Last and most important, I would like to thank Sophia for her
encouragement, support, patience and understanding that were
so valuable in every stage of this work and made completing
it possible.
% PC: whos was more patient - Sophia or Predrag?
