% Siminos thesis summary
% $Author: siminos $ $Date: 2009-03-13 16:26:49 -0400 (Fri, 13 Mar 2009) $

When statistical assumptions do not hold and coherent
structures are present in spatially extended systems
such as fluid flows, flame fronts and field theories,
a dynamical description of turbulent phenomena becomes
necessary.
% Recent experimental and theoretical advances\rf{science04}
% support a dynamical vision of turbulence:
% For any finite  spatial resolution, a turbulent flow follows approximately for a finite time
% a pattern belonging to a { finite alphabet} of admissible patterns.
% The long term dynamics is a {walk through the space of these unstable patterns}.
% When statistical assumptions\rf{frisch} break down and large
% scale coherent structures are present, a dynamical description
% as a walk through a space of unstable patterns becomes natural.
% The question is how to characterize and classify such patterns?
% Here we follow the seminal Hopf paper\rf{hopf48}, and  visualize
% turbulence not as  a sequence of spatial snapshots in turbulent evolution,
% but as a trajectory in an  $\infty$-$d$ \statesp\ in which an
% instant in turbulent evolution is a {unique} point.
In the dynamical systems approach, theory of turbulence for a
given system, with given boundary conditions, is given by (a)
the geometry of its $\infty$-dimensional \statesp\ and (b) the associated measure,
\ie, the likelihood that asymptotic dynamics visits a given
\statesp\ region.

In this thesis this vision is pursued in the context of \KS\
system, one of the simplest physically interesting spatially
extended nonlinear systems.
With periodic boundary conditions, continuous
translational symmetry endows \statesp\ with additional
structure that often dictates the type of observed solutions.
At the same time, the notion of recurrence becomes relative:
asymptotic dynamics visits the neighborhood of any
equivalent, translated point, infinitely often.
Identification of points related by the symmetry group
action, termed symmetry reduction, although conceptually
simple as the group action is linear, is hard to implement in
practice, yet it leads to dramatic simplification of
dynamics.

Here we propose a scheme, based on the method of moving
frames of Cartan, to efficiently project solutions of
high-dimensional truncations of partial differential
equations computed in the original space to a reduced
\statesp. The procedure simplifies the visualization of
high-dimensional flows and provides new insight into the role
the unstable manifolds of equilibria and traveling waves play
in organizing Kuramoto-Sivashinsky flow. This in turn
elucidates the mechanism that creates unstable modulated
traveling waves (periodic orbits in reduced space) that
provide a skeleton of the dynamics. The compact
description of dynamics thus achieved sets the stage for
reduction of the dynamics to mappings between a set of
Poincar\'e sections.
